---
title: "Milk Classification"
author: "Alicia Key"
date: "2022-08-10"
output: html_document
---

# Milk Classification

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(readr)
library(dplyr)
library(rsample)
library(parsnip)
library(yardstick)
library(broom)
library(ggplot2)
library(tune)
```

## Load the data

```{r}
milk_raw <- read_csv("data/milknew.csv")
knitr::kable(head(milk_raw))
```

## Clean the data

Specifically, the outcome variable, `Grade`, needs to be a factor with two levels for use with yardstick at the end. The event of interest will be "good" milk. In the original dataset, there are 3 qualities of milk: high, medium, and low. I will group high and medium into "good" milk and low into "bad" milk.

Ultimately, the event of interest for the Quality column will be the first level of good milk.

```{r}
milk_clean <- milk_raw %>%
  transmute(
    pH,
    Temperature = Temprature,
    Odor,
    Fat,
    Turbidity,
    Color = Colour,
    Quality = factor(
      case_when(
        Grade == "high" ~ "good",
        Grade == "medium" ~ "good",
        Grade == "low" ~ "bad"
      ),
      levels = c("good", "bad")
    )
  )
```
